Handling String-Number Conversion in String Constraint Solving

### 2019-11-17 draft.tex
moving things around in progress

\newcommand{\nat}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
%\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\mathbb{P}}(#1)} %shouldn't it be something that takes automaton as parameter and returns formula?
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}
\newcommand{\encode}[1]{\mathit{encode}_{#1}}
\newcommand{\eqwrt}[1]{=_{#1}}
\newcommand{\iequiv}{\equiv}
\newcommand{\restrict}[2]{#1{\restriction_{#2}}}
\newcommand{\underf}[2]{\mathit{under}_#1(#2)}

@@

@@

The \emph{string-number conversion constraint} $\phi_s$ is of the form $n=\sti{x}$, where the function $\sti{x}$ is defined as follows. For $a\in [0,9]$, we have $\sti{a}=a$ and for $s \cdot a \in [0,9]^+$, $\sti{s\cdot a} = 10\times \sti{s}+a$. For $s\notin [0,9]^+$, $\sti{s}=-1$. We define  $\modelsof{\phi_s}= \{I \mid I(n)= \sti{I(x)} \}$.

A \emph{string constraint} is a constraint of the above form. We say that a set of string constraint $\Phi$ is \emph{satisfiable}  iff $\bigcap_{\phi \in \Phi} \modelsof{\phi}$ is not empty.

@@

\section{Under-approximation} \label{section:under_approximate}
In this section, we describe formally how to restrict the domain of string variables to patterns defined by \emph{parametric flat automata} (PFA) (Section~\ref{section:sfa}) and hence convert string constraints $\Phi_I$ to a set of equisatisfiable linear formulas $\Phi_O$ in polynomial-time. We assume all disequality constraint $t_1 \neq t2$ are already converted to equal-satisfiable equality constraints and integer constraints in the standard way~\cite{abdulla2015norn}.

We pick a PFA for each variable in $\vars$ at the beginning of the decision procedure and then construct the set of linear integer constraints in a stepwise manner. In each step, we take a string constraint from $\Phi_I$ and depending on the type of the string constraint, we execute the corresponding procedure to produce the linear integer constraints and add them to the output linear formulas $\Phi_O$. We also add constraints to $\Phi_O$ to ensure all character variables of the PFAs only pick values from $\Sigma_\epsilon$.

More concretely, at the beginning, we move all integer constraints in $\Phi_I$ to the output linear formulas $\Phi_O$. After this step, $\Phi_I$ is the union of the three sets of constraints: the equality constraints, the regular constraints, and the string-number conversion constraints. For all string variables $x$ that occurs in a string-number conversion constraint $n= \sti{x}$, we follow the rule that will be described in Section~\ref{section:s2i} to pick the corresponding PFA. We do not set any restriction to the selection of PFA for other string variables. The selection is an implementation decision. We will describe the PFA selection strategy of \tool in Section~\ref{section:implementation}.

After the PFA for each variable is selected, we step-wisely pick and remove a string constraint from $\Phi_I$, executing the corresponding procedure to create equisatisfiable linear constraints, and add all the results to $\Phi_O$. The procedures for regular, equality and string-number conversion constraints are in Sections~\ref{section:mem},~\ref{section:eq}, and~\ref{section:s2i}, respectively. In the rest of the section, we assume each string variable $x$ is restricted to a PFA $A_x$ with the set of character variables~$V_x$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Parametric Flat Automata} \label{section:sfa}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\st}[2]{q_{#1}^{#2}}
\newcommand{\sym}[2]{v_{#1}^{#2}}
\todo{additional presburget constraints to encode $\epsilon$ added at translation to a formula}
%We use $\cvar$ to denote a finite set of \emph{character variables}  ranging over $\Sigma_\epsilon$. 
%Sometimes we use index to denote different set of variables over the same range, e.g., $\cvarone$ and $\cvartwo$. 
%We use $v$ and its variants for character variables. We call a string over $\cvar$ (character variables) a \emph{symbolic string}.
%An \emph{interpretation $I$ over $\cvar$} is a mapping from $\cvar$ to $\Sigma_\epsilon$. For a symbolic string $x= v_1v_2\cdots v_k$ over $\cvar$, $I(x)$ is defined as $I(v_1)\cdot I(v_2)\cdot \ldots \cdot I(v_k)$.

We introduce \emph{parametric flat automata} (PFA) that will be used to define patterns 
used by the under-approximation module to restrict the domain of string variables. 
A \emph{parametric automaton} (PA) is a finite automaton operating over an alphabet $\cvar$ of \emph{character variables}. 
A parametric flat automaton is a parametric automaton $A = (Q,T,\cvar,\st{0}{0},\st{m}{0})$ satisfying the following structural constraints:
\begin{enumerate}
	\item The final state $\st{m}{0}$ is reached from  the initial state $\st{0}{0}$ through a straight path of $m-1$ transitions $(\st{i}{0},v_i,\st{i+1}{0}) \in T$, $\st{i}{0} \in Q$ for $i\in[0,m-1]$. %, with pairwise distiand states $\st{i}{0} \in Q$ for $i\in[0,m-1]$.
	\item  
Each state $\st{i}{0}$ is the origin of a unique simple cycle of the length $l_i\in\mathbb{N}$, consisting of states $\st{i}{j-1} \in Q$ and transitions $(\st{i}{j-1}, \sym{i}{j-1}, \st{i}{j \bmod l_i})$ for $j\in [1,l_i]$. 
Notice that case when $l_i = 0$ is also admissible and means that there is no cycle on $q_i$.
%Assume that $\st{i}{0}$ is included in a cycle with $l_i$ states (includes $\st{i}{0}$). 
%For each $j\in [1,l_i]$, we have $\st{i}{j} \in Q$ and $(\st{i}{j-1}, \sym{i}{j-1}, \st{i}{j \bmod l_i})$. 
%(where $l_i = 0$ means that there is no cycle on $q_i$)
	\item Each variable in $\cvar$ appears on at most one transition.% (but $\epsilon$ transition can appear multiple times). 
\end{enumerate} 
\begin{figure*}
	\tikzset{state/.style={circle,draw=blue!50,fill=blue!20,
			thick,inner sep=0pt,minimum size=6mm}, initial text=$ $}
	\begin{tikzpicture} 
		\node[state,initial] (q0)  {$\st{0}{0}$};
		\node[state] (q1) [right = 3cm of q0] {$\st{1}{0}$};
		\node (q2) [right = 1.8cm of q1]{};
		\node (q3) [right = 1cm of q2]{};
		\node[state,accepting] (qm) [right = 1.8cm of q3] {$\st{m}{0}$};

@@

%Such automata are recognizers of symbolic strings, i.e., strings over $\cvar$.
Parametric automata accept words over $\cvar$, called \emph{parametric words}, but we still use them as representations of languages over $\Sigma$. 
%
Namely, word over $\cvar$ are interpreted as words over $\Sigma$ given an assignment of characters to its character variables.  
%
Formally, an \emph{interpretation $I$ over $\cvar$} is a mapping from $\cvar$ to $\Sigma_\epsilon$ and for a parametric word $x= v_1v_2\cdots v_k$ over $\cvar$, $I(x)$ is defined as $I(v_1)\cdot I(v_2)\cdot \ldots \cdot I(v_k)$.
%
We define the semantics of $x$ as the set $\semof x$ of all possible interpretations of $x$, the semantics of a language $L\subseteq V^*$ is the union $\semof L = \cup_{x\in L} \semof{x}$ of semantics of its words, and the semantics $\semof A$ of an PFA is the semantics of its language.

The crucial feature of parametric flat automata is that their semantics can be faithfully represented by a linear formula and handled efficiently by an SMT solver. 
\lh{, allowing to circumvent more costly automata algorithms and decision procedures. Concretize and justify this better?}
%
This is facilitated by the structural constraints of PFA, which guarantee that 
every parametric word $x\in L(A)$ is uniquely determined by its Parikh image $\parikhof{x}$, i.e., $\decode A (\parikhof x)) = x$, and analogously for the entire language of $A$, as formalized by the following lemma.

\begin{lemma}\label{lemma:parikh}
For an PFA $A$ and $x,y\in L(A)$, if $\parikhof x = \parikhof y$ then $x = y$. 
\end{lemma}
%
Indeed, since every variable appears on at most one transition, then the $\parikhof x$ value of all variables appearing within the same cycle is the same, and it is equal to the number of repetitions of that cycle in the accepting run. 
%
The accepting run on $x$ (and so $x$ itself) can thus be reconstructed from $\parikhof x$. 
%
For example, in the automaton of Figure~\ref{fig:sfa_def}, 
from $\parikhwof x {v_0} = \parikhwof x {v_1} = \cdots = \parikhwof x {v_{m-1}}=1$ and $\parikhwof x {v^0_1}=\parikhwof x {v^1_1}=\parikhwof x {v^2_1}=2$ we derive that $x=v_0(v^0_1v^1_1v^2_1)^2v_1\cdots v_{m-1}$. 
We may say that $\parikhof x$ is an encoding of $x$, and define the decoding function $\decode A$ such that $\decode A(\parikhof x) = x$, and we lift the function to sets of integer assignments in the standard manner.
%

\begin{lemma}\label{lemma:decx}
%For a PFA $A$, $L(A) = \{\decode A(\pim) \mid \pim\in \parikhof A\}$.
For a PFA $A$, $L(A) = \decode A(\parikhof{L(A)}$.
\end{lemma}
%\lh{should we replace $\semof{\parikhfof A}$ by $\parikhof{A}$ here and almost everywhere?}
%\yfc{Is $\parikhof{A}$ defined only over $\#I$? Does this work for Corollary~\ref{corollary:pfa}?}
%
\begin{proof}
The function $\decode A$ can particularly be implemented as follows. Given $\pim:\pvars\rightarrow\nat$,
and assuming that the lengths of the loops of $A$ are $l_0,\ldots,l_{m-1}$, 
$\decode A(\pim)$ is constructed as the word $x_0 v_{0} \ldots v_{m-1} x_{m-1}$ where for each $i\in [0,m-1]$,
$x_i = (v_i^0 \cdots v_i^{l_{i-1}})^{\pvarsof {v_i^0}}$ if $l_i >0$ and $x_i = \epsilon$ if $l_i = 0$. 
\end{proof}

Since every word $w \in \semof{A}$ is obtained from a parametric word $x\in L(A)$ and an interpretation of character variables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric word $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different words are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup \parikhof x) = w$.
The following corrolary of Lemma~\ref{lemma:decx} states that encodings of words in $\semof{A}$ are encodings of parametric words of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$:%, which are precisely all models of the formula $\parikhfof A$ interpreted over $\cvar\cup\pvarsof \cvar$, as formulated by the following corollary of Lemma~\ref{lemma:decx}:
%
%\begin{corollary}\label{corollary:pfa}
%For an PFA $A$ over character variables $V$,
%$\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}$.
%%$\semof{A} = \{I(\decode A(\pim)) \mid \pim\in\semof{\parikhfof A},I:\cvars\rightarrow\Sigma_\epsilon\}$.
%\end{corollary}

\begin{corollary}\label{corollary:pfa}
For an PFA $A$ over character variables $V$,
%$\semof{A} = \{(\decode A(I\cup\pim)) \mid \pim\in\parikhof A,I:\cvars\rightarrow\Sigma_\epsilon\}$.
$\semof A = \decode A \{(I\cup \pim)\mid \pim\in\parikhof{L(A)} ,I:\cvars\rightarrow\Sigma_\epsilon\}$
\end{corollary}
%
A minor technical difficulty is that we wish to work with pure arithmetic formulas, 
but we may also wish to specify when a character variable is assigned $\epsilon$, which is not a number. 
Hence, to allow purely arithmetic formulas to talk about the value $\epsilon$, we encode $\epsilon$ as $-1$.
In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulas constraining interpretations of character variables.

@@

>>>>

\newcommand{\nat}{\mathbb{N}}
\newcommand{\integers}{\mathbb{Z}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
%\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\mathbb{P}}(#1)} %shouldn't it be something that takes automaton as parameter and returns formula?
\newcommand{\pim}{I_{\#}}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}
\newcommand{\encode}[1]{\mathit{encode}_{#1}}
\newcommand{\eqwrt}[1]{=_{#1}}
\newcommand{\iequiv}{\equiv}
\newcommand{\restrict}[2]{#1{\restriction_{#2}}}
\newcommand{\underf}[2]{\mathit{under}_#1(#2)}
\newcommand{\pfaf}{\psi}

\newcommand{\leftA}{A^{\mathit{left}}}
\newcommand{\rightA}{A^{\mathit{right}}}
\newcommand{\leftV}{V^{\mathit{left}}}
\newcommand{\rightV}{V^{\mathit{right}}}
\newcommand\renvars{V_R^\mathsf{ver}}

\newcommand\insc{\phi_{\mathit{in}}}
\newcommand\outlf{\Psi_{\mathit{out}}}

@@

We will use four types of \emph{atomic string constraint}:

@@

A \emph{string-number conversion constraint} $\phi_s$ is of the form $n=\sti{x}$, where the function $\sti{x}$ is defined as follows. For $a\in [0,9]$, we have $\sti{a}=a$ and for $s \cdot a \in [0,9]^+$, $\sti{s\cdot a} = 10\times \sti{s}+a$. For $s\notin [0,9]^+$, $\sti{s}=-1$. We define  $\modelsof{\phi_s}= \{I \mid I(n)= \sti{I(x)} \}$.

A \emph{string constraint} is then a conjunction of atomic string constraints, with the semantics defined in the standard manner. It is \emph{satisfiable} if its semantics is nonempty.

@@

\section{Under-approximation} \label{section:under_approximate}
In this section, we describe formally how to restrict the domain of string variables to patterns defined by \emph{parametric flat automata} (PFA) (Section~\ref{section:sfa}) and hence convert string constraints $\insc$ to a set of equisatisfiable linear formulas $\outlf$ in polynomial-time. We assume all disequality constraint $t_1 \neq t2$ are already converted to equal-satisfiable equality constraints and integer constraints in the standard way~\cite{abdulla2015norn}.

We pick a PFA for each variable in $\vars$ at the beginning of the decision procedure and then construct the set of linear integer constraints in a stepwise manner. In each step, we take a string constraint from $\insc$ and depending on the type of the string constraint, we execute the corresponding procedure to produce the linear integer constraints and add them to the output linear formulas $\outlf$. We also add constraints to $\outlf$ to ensure all character variables of the PFAs only pick values from $\Sigma_\epsilon$.

More concretely, at the beginning, we move all integer constraints in $\insc$ to the output linear formulas $\outlf$. After this step, $\insc$ is the union of the three sets of constraints: the equality constraints, the regular constraints, and the string-number conversion constraints. For all string variables $x$ that occurs in a string-number conversion constraint $n= \sti{x}$, we follow the rule that will be described in Section~\ref{section:s2i} to pick the corresponding PFA. We do not set any restriction to the selection of PFA for other string variables. The selection is an implementation decision. We will describe the PFA selection strategy of \tool in Section~\ref{section:implementation}.

After the PFA for each variable is selected, we step-wisely pick and remove a string constraint from $\insc$, executing the corresponding procedure to create equisatisfiable linear constraints, and add all the results to $\outlf$. The procedures for regular, equality and string-number conversion constraints are in Sections~\ref{section:mem},~\ref{section:eq}, and~\ref{section:s2i}, respectively. In the rest of the section, we assume each string variable $x$ is restricted to a PFA $A_x$ with the set of character variables~$V_x$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parametric Flat Automata} \label{section:sfa}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\st}[2]{q_{#1}^{#2}}
\newcommand{\sym}[2]{v_{#1}^{#2}}
\todo{additional presburget constraints to encode $\epsilon$ added at translation to a formula}
%We use $\cvar$ to denote a finite set of \emph{character variables}  ranging over $\Sigma_\epsilon$. 
%Sometimes we use index to denote different set of variables over the same range, e.g., $\cvarone$ and $\cvartwo$. 
%We use $v$ and its variants for character variables. We call a string over $\cvar$ (character variables) a \emph{symbolic string}.
%An \emph{interpretation $I$ over $\cvar$} is a mapping from $\cvar$ to $\Sigma_\epsilon$. For a symbolic string $x= v_1v_2\cdots v_k$ over $\cvar$, $I(x)$ is defined as $I(v_1)\cdot I(v_2)\cdot \ldots \cdot I(v_k)$.

We introduce \emph{parametric flat automata} that will be used to define patterns 
used by the under-approximation module to restrict the domain of string variables.

\paragraph{Flat automata}
First, we say that an automaton FA 
$A = (Q,T,\Sigma,\st{0}{0},\st{m}{0})$ is \emph{flat} if it satisfies the following structural constraints:
\begin{enumerate}
	\item The final state $\st{m}{0}$ is reached from  the initial state $\st{0}{0}$ through a straight path of $m-1$ transitions $(\st{i}{0},a_i,\st{i+1}{0}) \in T$, $\st{i}{0} \in Q$ for $i\in[0,m-1]$. 
	\item  
Each state $\st{i}{0}$ is the origin of a unique simple cycle of the length $l_i\in\mathbb{N}$, consisting of states $\st{i}{j-1} \in Q$ and transitions $(\st{i}{j-1}, a_{i}^{j-1}, \st{i}{j \bmod l_i})$ for $j\in [1,l_i]$. 
Notice that the case when $l_i = 0$ is also admissible and means that there is no cycle on $q_i$.
	\item Each character in $\Sigma$ appears on at most one transition. 
\end{enumerate}

The crucial feature of flat automata is that their semantics can be faithfully represented by a linear formula and handled efficiently by an SMT solver. 
\lh{, allowing to circumvent more costly automata algorithms and decision procedures. Concretize and justify this better?}
%
This is facilitated by flatness, which guarantees that 
every parametric word $w\in L(A)$ is uniquely determined by its Parikh image $\parikhof{w}$, i.e., $\decode A (\parikhof w)) = w$, and analogously for the entire language of $A$, as formalized by the following lemma.

\begin{lemma}\label{lemma:parikh}
For an flat FA $A$ and $w,w'\in L(A)$, if $\parikhof w = \parikhof {w'}$ then $w = w'$. 
\end{lemma}
%
Indeed, since every variable appears on at most one transition, then the $\parikhof w$ value of all variables appearing within the same cycle is the same, and it is equal to the number of repetitions of that cycle in the accepting run. 
%
The accepting run on $w$ (and so $w$ itself) can thus be reconstructed from $\parikhof w$.

For example, in the automaton of Figure~\ref{fig:sfa_def}, 
from $\parikhwof x {v_0} = \parikhwof x {v_1} = \cdots = \parikhwof x {v_{m-1}}=1$ and $\parikhwof x {v^0_1}=\parikhwof x {v^1_1}=\parikhwof x {v^2_1}=2$ we derive that $x=v_0(v^0_1v^1_1v^2_1)^2v_1\cdots v_{m-1}$.

We may say that $\parikhof w$ is an encoding of $w$, and define the decoding function $\decode A$ such that $\decode A(\parikhof w) = w$, and we lift the function to sets of integer assignments in the standard manner.
%

\begin{lemma}\label{lemma:decx}
%For a PFA $A$, $L(A) = \{\decode A(\pim) \mid \pim\in \parikhof A\}$.
For a flat FA $A$, $L(A) = \decode A(\parikhof{L(A)}$.
\end{lemma}
%\lh{should we replace $\semof{\parikhfof A}$ by $\parikhof{A}$ here and almost everywhere?}
%\yfc{Is $\parikhof{A}$ defined only over $\#I$? Does this work for Corollary~\ref{corollary:pfa}?}
%
\begin{proof}
The function $\decode A$ can particularly be implemented as follows. Given $\pim:\pvars\rightarrow\nat$,
and assuming that the lengths of the loops of $A$ are $l_0,\ldots,l_{m-1}$, 
$\decode A(\pim)$ is constructed as the word $w_0 a_{0} \ldots a_{m-1} w_{m-1}$ where for each $i\in [0,m-1]$,
$w_i = (a_i^0 \cdots a_i^{l_{i-1}})^{\pvarsof {a_i^0}}$ if $l_i >0$ and $w_i = \epsilon$ if $l_i = 0$. 
\end{proof}

\paragraph{Parametric (flat) automata}
Next we define \emph{parametric automaton} (PA) as a tuple $A = (Q,T,\cvar,\st{0}{0},\st{m}{0},\pfaf)$ where $\pfaf$ is an \emph{interpretation constraint}, a linear formula over $\cvar$, and $(Q,T,\cvar,\st{0}{0},\st{m}{0})$ is an automaton operating over an alphabet $\cvar$ of \emph{character variables}.  \emph{Parametric flat automaton (PFA)} then is a parametric automaton with its finite automaton flat.

\begin{figure*}
	\tikzset{state/.style={circle,draw=blue!50,fill=blue!20,
			thick,inner sep=0pt,minimum size=6mm}, initial text=$ $}
	\begin{tikzpicture} 
		\node[state,initial] (q0)  {$\st{0}{0}$};
		\node[state] (q1) [right = 3cm of q0] {$\st{1}{0}$};
		\node (q2) [right = 1.8cm of q1]{};
		\node (q3) [right = 1cm of q2]{};
		\node[state,accepting] (qm) [right = 1.8cm of q3] {$\st{m}{0}$};

@@

%Such automata are recognizers of symbolic strings, i.e., strings over $\cvar$.
Parametric automata accept words over $\cvar$, called \emph{parametric words}, but we still use them as representations of languages over $\Sigma$. 
%
Namely, words over $\cvar$ are interpreted as words over $\Sigma$ under an
\emph{interpretation of $\cvar$}
%
a mapping from $I:\cvar\rightarrow$ to $\Sigma_\epsilon$. For a parametric word $x= v_1v_2\cdots v_k$ over $\cvar$, $I(x)$ is defined as $I(v_1)\cdot I(v_2)\cdot \ldots \cdot I(v_k)$.
%
We define the semantics of $x$ as the set $\semof x$ of all possible interpretations of $x$, the semantics of a language $L\subseteq V^*$ is the union $\semof L = \cup_{x\in L} \semof{x}$ of semantics of its words, and the semantics of an PFA is the semantics of its language restricted by the formula $\pfaf$, that is,  $\semof A = \semof{L(A)}\cap\semof\pfaf$.

Since every word $w \in \semof{A}$ is obtained from a parametric word $x\in L(A)$ and an interpretation of character variables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric word $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different words are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup \parikhof x) = w$, and we again lift the decoding function into sets of interpretations in the standard manner.
The following corrolary of Lemma~\ref{lemma:decx} states that encodings of words in $\semof{A}$ are encodings of parametric words of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$ that satisfy $\pfaf$:%, which are precisely all models of the formula $\parikhfof A$ interpreted over $\cvar\cup\pvarsof \cvar$, as formulated by the following corollary of Lemma~\ref{lemma:decx}:
%
%\begin{corollary}\label{corollary:pfa}
%For an PFA $A$ over character variables $V$,
%$\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}$.
%%$\semof{A} = \{I(\decode A(\pim)) \mid \pim\in\semof{\parikhfof A},I:\cvars\rightarrow\Sigma_\epsilon\}$.
%\end{corollary}

\begin{corollary}\label{corollary:pfa}
For an PFA $A = (Q,T,\cvarone,q_i,q_f,\pfaf)$,
%$\semof{A} = \{(\decode A(I\cup\pim)) \mid \pim\in\parikhof A,I:\cvars\rightarrow\Sigma_\epsilon\}$.
\\
$\semof A = \decode A \{(I\cup \pim)\mid \pim\in\parikhof{L(A)} ,I\in\semof\pfaf\}$.
\end{corollary}
%

Last, we define the encoding function as a counterpart of the decoding. Given a PFA $A$, and a language $L\subseteq \Sigma^*$, we define the set of $A$-encodings of the words in $L$ as 
$\encode A (L) = \{e \mid \decode A (e) \in L\}$. 
%
Notice that only the words from $L\cap \semof A$ have an $A$-encoding, 
hence $\encode A(L) = \encode A(L\cap\semof A)$, and so due to Corollary~\ref{corollary:pfa}, $\decode A({\encode A}(L)) = L\cap\semof A$.
%$\encode A (L) = \{(I\cup\pim) \mid \decode A (I\cup \pim) \in L \cap \semof{A}\}$. 
%We have that $\decode A \circ \encode A$ equal the identity function.
%

\section{Flat Domain Restriction}
Let us now give a precise definition of how we restrict the domain of string constraints using flat automata. 
A \emph{flat domain restriction} for a string contraint $\insc$ over variables $X$ is a mapping $R$ that assigns to every variable $x\in X$ a PFA $R(x)$. 
For technical reasons that will be apparent further in this section, we require that if $x\neq  y$, then $R(x)$ and $R(y)$ operate over disjoint sets of character variables. We will use $V_R$ to denote the union of character variables of all $R(x),x\in X$.

The semantics of a string constraint $\phi$ restricted by $R$ is then defined as 
$\semof{\phi}^R = \{I\in\semof \phi \mid \forall x\in X:I(x)\in\semof{R(x)}\}$. 
In the rest of this section, we will show how to build an \emph{underapproximating formula} representing the semantics $\semof{\insc}^R$ of an input string constraint, that is, a formula $\outlf$ such that 
$\decode R (\semof {\outlf}) = \semof{\insc}_R$.

The formula $\outlf$ is built intuctively to the structure of $\insc$.
Namely, in the followin subsections we show how to build a formula $\underf R {\phi}$ for every atomic string constraint $\phi$, 
and for a conjunction of 
string constraints we have $\underf R {\phi \land \phi'} \defeq \underf R {\phi} \land \underf R {\phi'}$.
% and $\underf R {\neg\phi} \defeq \neg\underf R{\phi}$. The underapproximation formula is then in the form $\outlf \defeq \underf R {\insc}$.

The correctness of this construction, that is, that 
$$\decode R (\semof {\outlf}) = \semof{\insc}^R$$ can be proved by induction to the structure of $\insc$. However, for the induction step to go through, we will need a more meticulous view into the correspondence of $\phi$ and $\underf R \phi$. Particularly, we will need to
%pay attention to how $\underf R \phi$ atually encodes the models of $\phi$, and
 ensure that $\underf R \phi$ captures exactly all $R$-encodings of $\semof \phi^R$
(indeed, notice that if it would be allowed to capture only some of the encodings, then for instance $\underf R {\phi} \land \underf R {\phi'}$ could only underapproximate $\semof{\phi\land\phi'}^R$).

To this end, we define an additional notation.
For an interpretation $I:X\rightarrow \Sigma^*$, we define its $R$-encoding as the set $\encode R(I) = \bigcup_{x\in X} \encode A(I(x))$. We lift this notation to sets of interpretation in the standard manner, i.e., $\encode R (J) = \bigcup_{I \in J}\encode R (I)$ for a set of interpretations $J$.
%
The inductive argument needed in the proof of the correctness of the underapproximation then reads as
$$\semof{\underf R \phi}_{V_R\cup V_{\#R}}  = \encode R{\semof{\phi}}$$ 
We note that the restriction of the semantics of $\underf R \phi$ here is needed here since it is constructed with some additional variables.

\paragraph{Encoding $\epsilon$}
A minor technical difficulty is that we wish to build underapproximation in the form of pure arithmetic formulas, 
but we may also wish to specify when a character variable is assigned $\epsilon$, which is not a number. 
To allow purely arithmetic formulas to talk about $\epsilon$, we encode $\epsilon$ as $-1$.
In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulas constraining interpretations of character variables.

%\subsection{Handling Regular Constraints Symbolically} \label{section:mem}
\section{Flattening of basic string constraints}
By basic string constraints, we mean regular, equality, and lenght constraints.
First, we discuss operations over PA needed to implement them, synchronisation and concatenation.

@@

\paragraph{Concatenation of PFAs}
Concatenation is another operation over PFAs needed when constructing underapproximations. Its implementation is straightforward, connect the final state of the first PFA with the initial state of the second by an $\epsilon$-transition. Sinco our automata do not allow $\epsilon$-transition, 
the connecting transition is labeled by a fres variable $v_\epsilon$ and the constraint $v_\epsilon$ is added to the interpretation contraint.

Formally, given PFA $A = (Q,T,\cvarone,q_i,q_f,\pfaf)$ and ${A'} = (Q',T',\cvartwo,q_i',q_f',\pfaf')$ with $Q\cap Q'=\emptyset = \cvarone\cap\cvartwo$, 
their concatenation is the PFA $A\cdot A' = (Q\cup Q',T\cup T'\cup \{(q_f,v_\epsilon,q_i')\},\cvarone\cup\cvartwo\cup\{v_\epsilon\},q_i,q_f',\pfaf\land\pfaf'\land v_\epsilon = \epsilon)$ where $v_\epsilon$ is fresh, not from $\cvarone\cup\cvartwo$.
